---
title: "Problem Set 1"
author: "Raneem Rahman"
subtitle: MGSC 310 Problem Set Template
output:
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

library(knitr)

# Change the number in set seed to your own favorite number
set.seed(400)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = FALSE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```


<!-- ### start answering your problem set here -->
<!-- You may export your homework in either html or pdf, with the former usually being easier. 
     To export or compile your Rmd file: click above on 'Knit' then 'Knit to HTML' -->
<!-- Be sure to submit both your .Rmd file and the compiled .html or .pdf file for full credit -->


```{r setup-3}

# load all your libraries in this chunk 
library('tidyverse')
# note, do not run install.packages() inside a code chunk. install them in the console outside of a code chunk. 

```



## Question 1

1a) Text response to part a. 

```{r}

#  answered by submitting compiled markdown

```


1b) Response to part b. 

```{r}

#  answered by submitting compiled markdown

```

etc...


## Question 2

2a) Response to part a. 


```{r}
#2a
getwd()
# getwd() calls the current working directory

#2b
setwd("/Users/raneemrahman/Desktop/MGSC_310/datasets")
movies <- read_csv("IMDB_movies.csv")

#2c
num_rows <- nrow(movies)
num_columns <- ncol(movies)
dimensions <- dim(movies)
num_rows
num_columns
dimensions

#2d
column_names <- names(movies)
column_names

#2e
str(movies) #shows datatypes of each variable
# character variables: movie_title, director_name, country, language, genres, content_rating, color, plot_keywords, actor_1_name, actor_2_name

#2f
movies %>% slice(1:20)

#2g
top_10 <- movies %>% 
  arrange(desc(budget)) %>% 
  slice(1:10) %>% 
  select(director_name)
print(top_10)
  

2b) Response to part b. 
# Suppose you are a data scientist hired by a large clothing retailer in the US. Think of one causal and one predictive analysis that you could implement that could help clothing stores navigate the pressure of rising inflation.
# causal: changing the prices ( if the retailer were to do xyz, what would happen to rising inflation)
# predictive: model forecasts using previous data (if i can observe past data about retail, I can make predictions about rising inflation)


```{r}


```


## Question 3

Here is only text for question 3. 

Suppose you are a data scientist hired by a large clothing retailer in the US. Think of one causal and one predictive analysis that you could implement that could help clothing stores navigate the pressure of rising inflation.
causal: changing the prices ( if the retailer were to do xyz, what would happen to rising inflation)
predictive: model forecasts using previous data (if i can observe past data about retail, I can make predictions about rising inflation)

If you want to add a hypertext link [you can do so](http://www.chapman.edu)

- If you want to add bullets
  - You
  - Can 
  - Do 
  - It

If you want to add an image [follow this guide](https://stackoverflow.com/questions/25166624/insert-picture-table-in-r-markdown)



